\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-04-19 forwards-and-backwards.tex
.

Putting these together,
the level $u$ of an individual at $x$
evolves according to
$$
\dot{u}
=
    - \frac{v}{\lambda} \left(\lambda - v\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) 
    +
    \mu(x,\eta) u .
$$
We shall write 
$$
    b(x, \eta)
    :=
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    -
    \mu(x,\eta),
$$
which captures the local net difference between reproduction and death,
i.e., the deviation from criticality of the branching mechanism.
(This differs from $F = r\gamma - \mu$ defined above
because $b$ contains the adjustment to fecundity obtained by
averaging over establishment probabilities in the locations that offspring disperse to.)
We can then rewrite the differential equation 
governing the how each individual's level evolves as
\begin{align}
\dot{u}
    &=
    \lambda \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    \left\{
        \left(1-\frac{u}{\lambda}\right)^{2}
        - 1
        + 2 \frac{u}{\lambda}
    \right\}
    -
    b(x,\eta)u
    \nonumber \\
    &=
    \lambda^{-1} \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy) u^2
    -
    b(x, \eta)u
    . \label{differential equation for level}
\end{align}

@@

First consider the terms in the generator that come from birth events.
Recall that when a birth successfully establishes,
a new level is generated above the parent's level,
and this new level is assigned to either the offspring or the parent.
Since the probability of each is 1/2,
the contribution of birth to the generator is
\begin{align*}
\begin{split}
    f(\lp)
    \mapsto
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \frac{\theta}{\lambda} \gamma(x, \eta)
    \int_u^\lambda
    \int_{\IR^d}
    &
    \frac{1}{2}
    \bigg\{
        \left(
            g(y, u_1)
            -
            1
        \right)  \\
    &\qquad {}
        + \frac{ g(y, u) g(x, u_1) }{ g(x, u) }
    \bigg\}
    r(y, \eta) q(x, dy)
\end{split} .
\end{align*}
Here, $u_1$ is the new level and $y$ is the offspring's location,
and so the two terms in the integral correspond to the two situations:
in the first, we have added an individual at $(y, u)$,
while in the second, we replace an individual at $(x, u)$
by one at $(x, u_1)$ and another at $(y, u)$.

\comment{PLR edited through to here.}

>>>>

Putting these together,
the level $u$ of an individual at $x$
evolves according to
$$
\dot{u}
=
    - \frac{u}{\lambda} \left(\lambda - u\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q(x, dy) 
    +
    \mu(x,\eta) u .
$$
We shall write 
$$
    b(x, \eta)
    :=
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    -
    \mu(x,\eta),
$$
which captures the local net difference between reproduction and death,
i.e., the deviation from criticality of the branching mechanism.
(This differs from $F = \theta(r\gamma - \mu)$ defined above
both by the factor of $\theta$ and
because $b$ contains the adjustment to fecundity obtained by
averaging over establishment probabilities in the locations that offspring disperse to.)
We can then rewrite the differential equation 
governing the how each individual's level evolves as
\begin{align}
\dot{u}
    &=
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy)
    \left\{
        -\frac{u}{\lambda}\left(\lambda - u\right)
        + u
    \right\}
    -
    b(x,\eta)u
    \nonumber \\
    &=
    \lambda^{-1} \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy) u^2
    -
    b(x, \eta)u
    . \label{differential equation for level}
\end{align}

@@

First consider the terms in the generator that come from birth events.
Recall that when a birth successfully establishes,
a new level is generated above the parent's level,
and this new level is assigned to either the offspring or the parent.
Since the probability of each is 1/2,
the contribution of birth to the generator is
\begin{align*}
f(\lp)
&\mapsto
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \frac{\theta}{\lambda} \gamma(x, \eta)
    \int_u^\lambda
    \int_{\IR^d}
    \left(
    \frac{1}{2}
    \bigg\{
            g(y, u_1)
        + \frac{ g(y, u) g(x, u_1) }{ g(x, u) }
    \bigg\}
        - 1
    \right)
    r(y, \eta) q(x, dy)
    \\ \begin{split}
&=
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \gamma(x, \eta)
    \bigg\{
        \frac{1}{2 \lambda}
        \int_u^\lambda
        g(x, u_1) du_1
        \frac{
            \theta \int_{\IR^d} (g(y, u) - g(x, u)) r(y, \eta) q(x, dy)
        }{
            g(x, u)
        }
    \\ & \qquad \qquad \qquad {}
        + \frac{\theta}{\lambda}
        \int_u^\lambda \int_{\IR^d}
        \left( \frac{g(y, u_1) + g(x, v_1)}{2} - 1 \right)
        r(y, \eta) q(x, dy)
    \bigg\}
    .
    \end{split}
\end{align*}
Here, $u_1$ is the new level and $y$ is the offspring's location,
and so the two terms in the integral correspond to the two situations:
in the first, we have added an individual at $(y, u)$,
while in the second, we replace an individual at $(x, u)$
by one at $(x, u_1)$ and another at $(y, u)$.

The remaining term in the generator is due to the motion of particles' levels:
\begin{align*}
    f(\lp)
    \mapsto
    f(\lp)
    \theta
    \sum_{(x, u) \in \lp}
    \left(
        \lambda^{-1} \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q(x, dy) u^2
        -
        b(x, \eta)u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)} .
\end{align*}
